Online {\IterComp}

### 2017-09-16 results.tex
done

In this experiment we evaluate the runtime overhead introduced by the work efficiency profiling. We compile the program using the
\texttt{-O3}. We then measure the wall-clock time of the instrumented and the native versions to calculate the runtime overhead. For
each benchmark, we compute the average overhead over all its 1,000 inputs.
Figure~\ref{fig:overhead-O3} shows the runtime overhead imposed by the work profiling using each of the proposed work profiling strategies over
their non-instrumented counterparts.
%The WPO relaxation was performed using profiling information for every specific input, which provides
%the perfect information to perform the WPO relaxation. This means that the experiments show its best performance enabled by having perfect
%profiling information.

@@

While the \OptProf has a maximum overhead of near 60\%, the \WCRelax-\textit{5}\% has an overhead of
less than 16\% for all benchmarks, with the \WPRelax-\textit{5}\% reaches an average overhead of only 5\%,
compared to an average overhead of 12\% for the optimal profiling.

@@

The most critical case, amongst the evaluated benchmarks, happens with the
\texttt{adpcm\_d} program, with an overhead of about 59\% for the optimal
instrumentation.
With a threshold of only 5\%, the \WCRelaxLower relaxation reduces
the overhead by about 75\% over the \OptProf,
and the \WPRelax-5\% reduces it by about 80\%.
This benchmark has a single hot function
which consists mainly of a single hot loop with several branches inside it. %, as depicted by Figure~\ref{fig:adpcm_d-cfg-instr}.
Most of the improvement achieved by both strategies comes from removing only two probes from the hot loop. The two relaxed probes were
placed in branches with a high probability of being taken, but with a small contribution to the total amount of
work measured in DAG for this loop (both probes with an static error of about 1.3\%). Because these probes were placed in basic blocks with
a considerable execution frequency, removing the probes resulted in a significant reduction in runtime overhead.

A few of the benchmarks present abnormal regressions in their execution time.
This increase in the profiling overhead is counter-intuitive because the
relaxation algorithms only reduces the number of instrumented probes,
without changing their placement.
We suggest some reasons for these abnormal regressions.
By analyzing these cases, we noticed that some of the frequently executed
probes could not be removed by the relaxation strategies with small thresholds.
Another important aspect to point out is that we also observed
differences among the final CFGs after applying the \texttt{-O3} optimization.
Although relaxation does not directly alter the CFG, it is worth mentioning that
there are optimizations that make use of LLVM's built-in target-based cost model,
which means that adding or removing instructions from basic blocks may affect
decisions during transformations of the code, which may end up changing the CFG
in different ways depending on the instrumentation applied.
%For example, some of the optimizations that make use of this built-in cost model
%and also alter the CFG are: loop-unrolling, function inlining, and
%simplifications of the CFG.
%These changes in the program's CFGs may affect other compiler optimizations and
%also possibly in hardware performance, for example affecting the instruction
%caching.

Finally, as shown in Figure~\ref{fig:error-O3}, these improvements can be
obtained while incurring very little dynamic error.
This result confirms the overly conservative aspect of the \WCRelaxLower relaxation,
while the \WPRelaxLower relaxation is significantly less conservative.
Although the error bound is not guaranteed by the \WPRelaxLower relaxation,
its dynamic error was always below the 5\% thresholds.
%due to the use of perfect profiling information.

@@

In all configurations, the same optimization sequence is used for multiple inputs, using a dynamic input-window size, as explained in
Section~\ref{sec:oic-infra}. The average performance over the input window provides an estimate for the overall performance of the
optimization sequence across distinct inputs. The optimization sequences are ranked based on their average performance, and the best
optimization sequence is selected.

>>>>

In this experiment we evaluate the runtime overhead introduced by the work efficiency profiling. We compile the program using the
\texttt{-O3}. We then measure the wall-clock time of the instrumented and the native versions to calculate the runtime overhead. For
each benchmark, we compute the average overhead over all its 1,000 inputs. 
Figure~\ref{fig:overhead-O3} shows the runtime overhead introduced by profiling for each of the proposed strategies.
%The WPO relaxation was performed using profiling information for every specific input, which provides
%the perfect information to perform the WPO relaxation. This means that the experiments show its best performance enabled by having perfect
%profiling information.

@@

The overhead for \OptProf typically remains below or around 10\% but can reach up to 59\% for \texttt{adpcm\_d}, which would be
unacceptable. \WCRelax-\textit{5}\% on the other hand manages to eliminate most of the overhead, with a worst
case of 16\% and an average of 8\%. \WPRelax-\textit{5}\% further reduces the cost of profiling, 5\% overhead on average and 13\%
worst case. For both \WCRelax and \WPRelax setting the threshold to 2\% makes it too difficult to remove probes. As a result, overhead
remain high.

@@

The benchmark where \OptProf records its worst result, \texttt{adpcm\_d}, has a single hot function which consists mainly of a single hot
loop with several branches. This leads to numerous short basic blocks, many of them with probes. Most of the improvement achieved by both
relaxation strategies comes from removing only two probes from the hot loop. They were initially placed in branches with a high probability
of being taken but with a small work contribution. Removing either of the probes introduces error of about 1.3\% at the loop level but
eliminates three quarters of the overhead.

For a few of the benchmarks, relaxation unexpectedly introduces additional overhead. This is counter-intuitive since relaxation leads to
less computational work. In these cases, relaxation did not manage to remove high frequency probes, so we did not benefit significantly.
On the other, the probes that were removed affected how \texttt{-O3} optimized the surrounding code, in some cases blocking useful
optimizations that were applied originally. This resulted in slower code.
%For example, some of the optimizations that make use of this built-in cost model
%and also alter the CFG are: loop-unrolling, function inlining, and
%simplifications of the CFG.
%These changes in the program's CFGs may affect other compiler optimizations and
%also possibly in hardware performance, for example affecting the instruction
%caching.

In Figure~\ref{fig:error-O3}, we see the error in our work estimation introduced by relaxation. The error for the WC-Relax strategies is
too low, always below 1\% and usually much lower, regardless of threshold. This confirms that WC-Relax is too conservative. WP-Relax brings
us closer to our target threshold. Despite not having hard guarantees for the error under WP-Relax, in all cases the error was below the
threshold.

@@

For all strategies, the same optimization sequence is used for multiple inputs, using a dynamic input-window size, as explained in
Section~\ref{sec:oic-infra}. The average performance over the input window provides an estimate for the overall performance of the
optimization sequence across distinct inputs. The optimization sequences are ranked based on their average performance, and the best
optimization sequence is selected.
